---
title: "Bikes in DC"
output: html_notebook
---
The following is a tutorial originally from https://minecr.shinyapps.io/dsbox-06-dcbikeshare/. This tutorial involves viewing the affect of a variety of variables on bike ridership in Washington D.C.

###Packages
Load the packages we need for this tutorial using the chunk below
```{r}
library(tidyverse)
library(tidymodels)
library(dsbox)
library(readxl)
bikes <- read_excel("U:/DS241/202008-capitalbikeshare-tripdata.xlsx")

```

###Data
The data include daily bike rental counts (by members and casual users) of Capital Bikeshare in Washington, DC in 2011 and 2012 as well as weather information on these days. The data is contained in the dsbox package and is called dcbikeshare.

The original data sources are http://capitalbikeshare.com/system-data and http://www.freemeteo.com.

###Recode Data

##Recode Season
Recode the season variable to be a factor with meaningful level names as outlined in the codebook, with spring as the baseline level.
```{r}
dcbikeshare <- dcbikeshare %>%
  mutate(
    season = case_when(
      season == 1 ~ "winter",
      season == 2 ~ "spring",
      season == 3 ~ "summer",
      season == 4 ~ "fall"
    ),
    season = fct_relevel(season, "spring", "summer", "fall", "winter")
  )
```

##Recode holiday and workingday
Recode the binary variables holiday and workingday to be factors with levels no (0) and yes (1), with no as the baseline level.
```{r}
dcbikeshare <- dcbikeshare %>%
  mutate(
    holiday = ifelse(holiday == 0, "no", "yes"),      
    holiday = fct_relevel(holiday, "no", "yes"),    
    workingday = ifelse(workingday == 0, "no", "yes"),
    workingday = fct_relevel(workingday, "no", "yes")
  )
```

##Recode year
Recode the yr variable to be a factor with levels 2011 and 2012, with 2011 as the baseline level.
```{r}
dcbikeshare <- dcbikeshare %>%
  mutate(
    yr = ifelse(yr == 0, "2011", "2012"),
    yr = fct_relevel(yr, "2011", "2012")
  )
```

##Recode weathersit
Recode the weathersit variable as 1 - clear, 2 - mist, 3 - light precipitation, and 4 - heavy precipitation, with clear as the baseline.
```{r}
dcbikeshare <- dcbikeshare %>%
  mutate(
    weathersit = case_when(
      weathersit == 1 ~ "clear",
      weathersit == 2 ~ "mist",
      weathersit == 3 ~ "light precipitation",
      weathersit == 4 ~ "heavy precipitation"
    ),
    weathersit = fct_relevel(weathersit, "clear", "mist", "light precipitation", "heavy precipitation")
  )
```

Recoding warning
At this point, you might wonder what this warning means:

`Warning: Unknown levels in `f`: heavy precipitation`
This is simply R warning us that while it is possible for the weathersit variable to have the value 4 for heavy precipitation, this never actually occurs in the data. Consequently, the weathersit factor only has three levels.

###Creating new variables

##Create raw values for temperature, humidity, and windspeed
Calculate raw temperature, feeling temperature, humidity, and windspeed as their values given in the dataset multiplied by the maximum raw values stated in the codebook for each variable. Instead of writing over the existing variables, create new ones called temperature_raw, feeling_temperature_raw, humidity_raw, windspeed_raw.

```{r}
dcbikeshare <- dcbikeshare %>%
  mutate(
    temperature_raw = temp * 41,
    feeling_temperature_raw = atemp * 50,
    humidity_raw = hum * 100,
    windspeed_raw = windspeed * 67
  )
```

##Check cnt variable
Check that the sum of casual and registered adds up to cnt for each record. Do this by creating a new column that takes on the value TRUE if they add up and FALSE if not, and then checking if all values in that column are TRUEs.
```{r}
dcbikeshare %>%
  mutate(cas_plus_reg = casual + registered) %>%
  summarise(all_zero = all(cas_plus_reg == cnt))
```

###Bike rentals and temperature

##Recreate Visualization
```{r}
dcbikeshare %>%
  ggplot(mapping = aes(x = dteday, y = cnt, color = feeling_temperature_raw)) +
    geom_point(alpha = 0.7) +
    labs(
      title = "Bike rentals in DC, 2011 and 2012",
      subtitle = "Warmer temperatures associated with more bike rentals",
      x = "Date",
      y = "Bike rentals",
      color = "Temperature (C)"
    ) +
  theme_minimal()
```

##Bike rentals vs temperature
Fit a linear model predicting total daily bike rentals from raw daily temperature and answer the questions below.
```{r}
cnt_tmp <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ temperature_raw, data = dcbikeshare)
cnt_tmp %>%
  tidy()
```

Use the chunk below to print out the R^2 of the model
```{r}
glance(cnt_tmp)$r.squared
```

##Bike rentals vs feeling temperature
Fit another linear model predicting total daily bike rentals from raw daily feeling temperature. Then, proceed to answering the questions below
```{r}
cnt_atmp <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ feeling_temperature_raw, data = dcbikeshare)
cnt_atmp %>%
  tidy()
```

Use the chunk below to print out the R2 of the model.
```{r}
glance(cnt_atmp)$r.squared
```

###Full model
Fit a model predicting total daily bike rentals from season, year, whether the day is holiday or not, whether the day is a workingday or not, the weather category, temperature, feeling temperature, humidity, and windspeed, as well as the interaction between feeling temperature and holiday.
```{r}
cnt_full <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season + yr + holiday + workingday + weathersit +
        temperature_raw + feeling_temperature_raw + humidity_raw +
        windspeed_raw + feeling_temperature_raw * holiday, 
      data = dcbikeshare)
tidy(cnt_full)
```

Record adjusted R2 of the model.
```{r}
glance(cnt_full)$adj.r.squared
```

###Backward selection
Perform the first step of backward selection by fitting a series of models, each with one explanatory variable removed from the full model you fit in the previous exercise. Record the adjusted R2s of each of these models.

In the chunk below, fit the full model without the season variable and print out the R2 score.
```{r}
rm_season <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ yr + holiday + workingday + weathersit +
        temperature_raw + feeling_temperature_raw + humidity_raw +
        windspeed_raw + feeling_temperature_raw * holiday, 
      data = dcbikeshare)
glance(rm_season)$adj.r.squared
```

In the chunk below, fit the full model without the year variable and print out the adjusted R2 score.
```{r}
rm_year <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season+holiday + workingday + weathersit +
        temperature_raw + feeling_temperature_raw + humidity_raw +
        windspeed_raw + feeling_temperature_raw * holiday, 
      data = dcbikeshare)
glance(rm_year)$adj.r.squared
```

In the chunk below, fit the full model without the holiday variable and print out the adjusted R2 score. Remember to also remove the interaction effect between holiday and feeling_temperature_raw.
```{r}
rm_holiday <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season + yr + workingday + weathersit +
        temperature_raw + feeling_temperature_raw + humidity_raw +
        windspeed_raw, 
      data = dcbikeshare)
glance(rm_holiday)$adj.r.squared
```

Fit the full model without the workingday variable using the code chunk below! Remember to record the adjusted R2 score.
```{r}
rm_workingday<- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season+holiday + yr + weathersit +
        temperature_raw + feeling_temperature_raw + humidity_raw +
        windspeed_raw + feeling_temperature_raw * holiday, 
      data = dcbikeshare)
glance(rm_workingday)$adj.r.squared

```

Fit the full model without the weathersit variable in the chunk below and record the adjusted R2 score.
```{r}
rm_weathersit<- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt~ season+yr + holiday + workingday +
        temperature_raw + feeling_temperature_raw + humidity_raw +
        windspeed_raw + feeling_temperature_raw * holiday, 
      data = dcbikeshare)
glance(rm_weathersit)$adj.r.squared
```

Remove the temperature_raw variable in the chunk below and record the adjusted R2 score
```{r}
rm_temp_raw<- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season+yr + holiday + workingday + weathersit +
        feeling_temperature_raw + humidity_raw +
        windspeed_raw + feeling_temperature_raw * holiday, 
      data = dcbikeshare)
glance(rm_temp_raw)$adj.r.squared
```

In the chunk below, fit the full model without the feeling_temperature_raw variable and record the adjusted R2 score. Don’t forget to also remove the interaction effect between holiday and feeling_temperature_raw!
```{r}
rm_feeltemp_raw <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season+yr + holiday + workingday + weathersit +
        temperature_raw + humidity_raw +
        windspeed_raw, 
      data = dcbikeshare)
glance(rm_feeltemp_raw)$adj.r.squared
```

Fit the full model without the humidity_raw variable and record the adjusted R2 score.
```{r}
rm_humidity_raw <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season+yr + holiday + workingday + weathersit +
        temperature_raw + feeling_temperature_raw  +
        windspeed_raw + feeling_temperature_raw * holiday, 
      data = dcbikeshare)
glance(rm_humidity_raw)$adj.r.squared
```

Fit the full model without the windspeed_raw variable and record the adjusted R2 score.
```{r}
rm_windspeed_raw <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season+yr + holiday + workingday + weathersit +
        temperature_raw + feeling_temperature_raw + humidity_raw +
         feeling_temperature_raw * holiday, 
      data = dcbikeshare)
glance(rm_windspeed_raw)$adj.r.squared
```

Last, remove the interaction effect between holiday and feeling_temperature_raw without removing the two variables individually. Then, record the adjusted R2.
```{r}
rm_interaction <- linear_reg() %>%
  set_engine("lm") %>%
  fit(cnt ~ season+yr + holiday + workingday + weathersit +
        temperature_raw + feeling_temperature_raw + humidity_raw +
        windspeed_raw, 
      data = dcbikeshare)
glance(rm_interaction)$adj.r.squared
```


```{r}
library(readxl)
df1 <- read_excel("U:/DS241/202008-capitalbikeshare-tripdata.xlsx")
```

###Ridership vs time
```{r}
df4=df1 %>% 
  select(rideable_type, start=started_at, end=ended_at) %>%
  pivot_longer(cols=start:end,names_to="type", values_to="time") %>%
  arrange(time) %>%
  mutate(inc=case_when(
    type=="start" ~ 1,
    type=="end" ~ -1
  ), riders=cumsum(inc))

```

Can we plot over the entire month
```{r}
df4 %>% mutate(day=day(time)) %>%
  ggplot(aes(time, riders))+geom_point(size=.5, alpha=.02)+facet_wrap(~day)

```
